### 2014-06-03 ACL2014/discussion.tex
spelling fixes

The other example is to remove duplicated instances in a result set. 
When a query set is retrieved from a repository and presented to users, it is important to remove similar queries as user experiences would be distracted from ``duplicated results.'' 
That is, given a set of query terms, we can apply  \dflipb algorithm to build a lookup table of near-duplicates 
to find ``duplicated query terms'' efficiently. As ``duplicateness'' of among query terms typically requires ``higher'' degree of similarity (relatively easier problem) than ``relatedness'', we can tune parameters ($K, L, F$) based on a specific $\tau$ (e.g $\tau=0.9$) from training samples.  The fourth column in Table~\ref{tab:lists} illustrates several duplicate examples -- ``trumbull weather ct'' and ``weather in trumbull ct''.

We first described the existing ``vanilla'' \lsh algorithm. It is the first paper according to 
best of our knowledge that applies this algorithm to NLP applications. 
Next, we poposed four novel variants of vanilla \lsh. Two of our variants achieve 
significantly better recall than the vanilla LSH by using the same number of hash tables. 
We present a framework on Hadoop that efficiently finds nearest neighbors for a given query from a commercial large-scale query logs (consisting of hundreds of millions of queries) in sublinear time.

>>>>

The other example is to remove duplicated instances in a result set. 
When a query set is retrieved from a repository and presented to users, it is important to remove similar queries as user experiences would be distracted from ``duplicated results.'' 
That is, given a set of query terms, we can apply  \dflipb algorithm to build a lookup table of near-duplicates 
to find ``duplicated query terms'' efficiently. As ``duplicates'' among query terms typically requires ``higher'' degree of similarity (relatively easier problem) than ``relatedness'', we can tune parameters ($K, L, F$) based on a specific $\tau$ (e.g $\tau=0.9$) from training samples.  The fourth column in Table~\ref{tab:lists} illustrates several duplicate examples -- ``trumbull weather ct'' and ``weather in trumbull ct''.

We first described the existing ``vanilla'' \lsh algorithm. It is the first paper according to 
best of our knowledge that applies this algorithm to NLP applications. 
Next, we proposed four novel variants of vanilla \lsh. Two of our variants achieve 
significantly better recall than the vanilla LSH by using the same number of hash tables. 
We present a framework on Hadoop that efficiently finds nearest neighbors for a given query from a commercial large-scale query logs (consisting of hundreds of millions of queries) in sublinear time.

### 2014-06-03 ACL2014/experiments.tex
spelling fixes

In the second experiment, we vary $L$ to be in $\{1,10,28,55\}$ while fixing $K=16$ on the \aol and \dataA datasets. 
Recal that $L$ denotes the number of hash tables and $K$ is the width of the index of the buckets in the table. 
Increasing $K$ results in increase of the precision by reducing the false positive candidates, but 
the $L$ also need to be correspondingly increased in order the maintain a good recall
(i.e. reduce false negatives). 
Table \ref{tab:varyL} shows that increasing $L$ leads to better recall, 
but at the expense of performing more comparisons on both the datasets. 
In addition, having a large $L$ 
means generating large number of random projection bits and hash tables which is 
both time and memory intensive. Hence, we fix $L=10$, 
which leads to a reasonable recall with a tolerable number of comparisons.

\begin{table}
\centering
\begin{tabular}{|c|rc|rc|}
\hline
\multirow{2}{*}{L} & \multicolumn{2}{c|}{\small \aol} & \multicolumn{2}{c|}{\small \dataA} \\
 & {\small Comparisons} & {\small Recall} & {\small Comparisons} & {\small Recall}  \\
\hline
1 & 7  & .28 & 106 & .36 \\
 \rowcolor[gray]{0.9} 10  &  57 & .63 & 1052 & .67 \\
28 &  152 & .77 & 2908 & .78 \\
55 &  297 & .89 & 5648 & .84 \\
\hline 
 \end{tabular}
\caption{\footnotesize{Varying $L$ with fixed $K=16$ on \aol and \dataA with $\tau=0.7$.}}
\label{tab:varyL}
\end{table}
In the third experiment, we vary $K$ to be in $\{4,8,16\}$ while fixing $L=10$ on \aol and \dataA datasets. 
As expected, Table \ref{tab:varyK} shows that increasing $K$ reduces the number of 
comparisons and worsens recall on both the datasets. This is intuitive as  
the larger value of $K$ leads to larger 
gap between probabilites of collision between close queries 
and far queries (see Section \ref{sec:vlsh}). 
Hence, we fix $K=16$ to have few number of comparisons.

@@

In the fourth experiment, Table \ref{tab:lists} shows the qualitative results for some arbitrary queries. These results are found by applying our
system (\dflipb with parameters $L=10$, $K=24$, and $F=2$ ) on \dataC. The second column in Table \ref{tab:lists} shows that the returned approximate 
similar neighbors can be useful in finding related queries \cite{Jones06WWW,Jain11SIGIR}. 
The third column shows an example, where we can find several popular spellling errors. 
The first and last column show examples of near-duplicate queries \cite{leeCIKM11}.

>>>>

In the second experiment, we vary $L$ to be in $\{1,10,28,55\}$ while fixing $K=16$ on the \aol and \dataA datasets. 
Recall that $L$ denotes the number of hash tables and $K$ is the width of the index of the buckets in the table. 
Increasing $K$ results in increase of the precision by reducing the false positive candidates, but 
the $L$ also need to be correspondingly increased in order the maintain a good recall
(i.e. reduce false negatives). 
Table \ref{tab:varyL} shows that increasing $L$ leads to better recall, 
but at the expense of performing more comparisons on both the datasets. 
In addition, having a large $L$ 
means generating large number of random projection bits and hash tables which is 
both time and memory intensive. Hence, we fix $L=10$, 
which leads to a reasonable recall with a tolerable number of comparisons.

\begin{table}
\centering
\begin{tabular}{|c|rc|rc|}
\hline
\multirow{2}{*}{L} & \multicolumn{2}{c|}{\small \aol} & \multicolumn{2}{c|}{\small \dataA} \\
 & {\small Comparisons} & {\small Recall} & {\small Comparisons} & {\small Recall}  \\
\hline
1 & 7  & .28 & 106 & .36 \\
 \rowcolor[gray]{0.9} 10  &  57 & .63 & 1052 & .67 \\
28 &  152 & .77 & 2908 & .78 \\
55 &  297 & .89 & 5648 & .84 \\
\hline 
 \end{tabular}
\caption{\footnotesize{Varying $L$ with fixed $K=16$ on \aol and \dataA with $\tau=0.7$.}}
\label{tab:varyL}
\end{table}
In the third experiment, we vary $K$ to be in $\{4,8,16\}$ while fixing $L=10$ on \aol and \dataA datasets. 
As expected, Table \ref{tab:varyK} shows that increasing $K$ reduces the number of 
comparisons and worsens recall on both the datasets. This is intuitive as  
the larger value of $K$ leads to larger 
gap between probabilities of collision between close queries 
and far queries (see Section \ref{sec:vlsh}). 
Hence, we fix $K=16$ to have few number of comparisons.

@@

In the fourth experiment, Table \ref{tab:lists} shows the qualitative results for some arbitrary queries. These results are found by applying our
system (\dflipb with parameters $L=10$, $K=24$, and $F=2$ ) on \dataC. The second column in Table \ref{tab:lists} shows that the returned approximate 
similar neighbors can be useful in finding related queries \cite{Jones06WWW,Jain11SIGIR}. 
The third column shows an example, where we can find several popular spelling errors. 
The first and last column show examples of near-duplicate queries \cite{leeCIKM11}.

### 2014-06-03 ACL2014/system.tex
Updated bin
spelling fixes

\subsection{Vanilla \lsh}
\label{sec:vlsh}
The \lsh algorithm relies on the existence of an family of locality sensitive hash functions. 
Let $H$ be a family of hash functions mapping 
 $\mathbb{R}^D$ to some universe S. For any two query terms $p$, $q$; we chose $h\in H$ uniformly at random; 
 and analyze the probability that $h(p) = h(q)$.  
 Suppose $d$ is a distance function (cosine distance for us), 
$R > 0$ be a distance threshold $R>0$ and $c>1$ be an approximation factor.   
Let $P_1, P_2 \in (0, 1)$ be two probability thresholds.  
The family $H$ of hash functions is called a $(R, cR, P_1, P_2)$ locality sensitive family if 
 it satisfies the following conditions:
\begin{enumerate}
\item If $d(p,q) \leq R$, then $Pr[h(p)=h(q)] \geq P_1$, 
\item and if $d(p,q) \geq cR$, then $Pr[h(p)=h(q)] \leq P_2$
\end{enumerate}
An \lsh family is generally interesting when  $P_1>P_2$. 
However, the difference between $P_1$ and $P_2$ can be very small. 
Given a family $H$ of hash functions with parameters $(R, cR, P_1,P_2)$, 
the \lsh algorithm is devised by amplifying the gap between 
the two probabilities $P_1$ and $P_2$ by concatenating several functions. 
In particular, \lsh algorithm concatenates $K$ hash functions
to create a new hash function $g(\cdot)$ as:
$g(q)=(h_1(q),h_2(q),\dots,h_K(q))$. 
A larger value of $K$ leads to larger 
gap between probabilites of collision between close neighbors (i.e. distance less than $R$) and neighbors that are far (i.e. 
distance more than $cR$); 
the corresponding probabilities being $P_1^{K}$ and $P_2^{K}$ respectively. 
This amplification ensures a high precision of the algorithm by  
making the probability of dissimilar queries having the same hash value very small.

@@

\begin{figure}[ht]
\small
\fbox{
\begin{minipage}{2.5in}
{\bf Preprocessing:} Input is $N$ queries with their respective feature vectors.
\begin{itemize}
\item Select $L$ functions $g_j$, $j=1,2,\dots,L$, setting  $g_j(q)=(h_{1,j}(q),h_{2,j}(q),\dots,h_{K,j}(q))$, where 
$\{h_{i,j}, i\in [1, K], j\in [1,L]\}$ are chosen at random from the \lsh family.
\item Construct $L$ hash tables, $\forall 1 \leq j \leq  L$. 
All queries with the same $g_j$ value ($\forall 1 \leq j \leq  L$) are placed in the same bucket.    
\end{itemize}

@@

\subsection{Multi Probe \lsh}
\label{sec:mblsh}
As we discussed in Section \ref{subsec:resuseHash}, generating hash functions can be computationally expensive.
Since the memory required by the algorithm also scales linearly with $L$, the number of hash tables, it is desirable to have a small 
number of tables to reduce the memory footprint. The memory footprint of vanilla LSH is what makes it impractical for real applications. 
Here, we first describe four new variants of the vanilla \lsh algorithm motivated by the intuition in
\mblshf~\cite{LvVLDB07}. \mblshf obtains significantly higher recall than the vanilla \lsh while using
the same number of hash tables. In order to achieve this, the main intuition utilized in \mblshf is that
in addition to looking at the hash bucket that a test query $q$ falls in, it is also possible to look at the neigboring buckets
in order to find its near neigbour candidates. \mblshf in ~\cite{LvVLDB07} suggests exploring the neighboring buckets in order of the
Hamming distance from the bucket in which $q$ falls. They then empirically show that these neighboring buckets contain 
the near neighbors with very high probability. 
Although \mblshf is achieves a higher recall for the same number of hash tables, it will also require more 
probes since it searches for multiple buckets within a table. 
The main advantage of searching in multiple buckets over generating more number of tables 
is that it takes more memory and time to generate more tables in pre-processing.

>>>>

\subsection{Vanilla \lsh}
\label{sec:vlsh}
The \lsh algorithm relies on the existence of an family of locality sensitive hash functions. 
Let $H$ be a family of hash functions mapping 
 $\mathbb{R}^D$ to some universe S. For any two query terms $p$, $q$; we chose $h\in H$ uniformly at random; 
 and analyze the probability that $h(p) = h(q)$.  
 Suppose $d$ is a distance function (cosine distance for us), 
$R > 0$ be a distance threshold $R>0$ and $c>1$ be an approximation factor.   
Let $P_1, P_2 \in (0, 1)$ be two probability thresholds.  
The family $H$ of hash functions is called a $(R, cR, P_1, P_2)$ locality sensitive family if 
 it satisfies the following conditions:
\begin{enumerate}
\item If $d(p,q) \leq R$, then $Pr[h(p)=h(q)] \geq P_1$, 
\item and if $d(p,q) \geq cR$, then $Pr[h(p)=h(q)] \leq P_2$
\end{enumerate}
An \lsh family is generally interesting when  $P_1>P_2$. 
However, the difference between $P_1$ and $P_2$ can be very small. 
Given a family $H$ of hash functions with parameters $(R, cR, P_1,P_2)$, 
the \lsh algorithm is devised by amplifying the gap between 
the two probabilities $P_1$ and $P_2$ by concatenating several functions. 
In particular, \lsh algorithm concatenates $K$ hash functions
to create a new hash function $g(\cdot)$ as:
$g(q)=(h_1(q),h_2(q),\dots,h_K(q))$. 
A larger value of $K$ leads to larger 
gap between probabilities of collision between close neighbors (i.e. distance less than $R$) and neighbors that are far (i.e. 
distance more than $cR$); 
the corresponding probabilities being $P_1^{K}$ and $P_2^{K}$ respectively. 
This amplification ensures a high precision of the algorithm by  
making the probability of dissimilar queries having the same hash value very small.

@@

\begin{figure}[ht]
%\small
\fbox{
\begin{minipage}{2.5in}
{\bf Preprocessing:} Input is $N$ queries with their respective feature vectors.
\begin{itemize}
\item Select $L$ functions $g_j$, $j=1,2,\dots,L$, setting  $g_j(q)=(h_{1,j}(q),h_{2,j}(q),\dots,h_{K,j}(q))$, where 
$\{h_{i,j}, i\in [1, K], j\in [1,L]\}$ are chosen at random from the \lsh family.
\item Construct $L$ hash tables, $\forall 1 \leq j \leq  L$. 
All queries with the same $g_j$ value ($\forall 1 \leq j \leq  L$) are placed in the same bucket.    
\end{itemize}

@@

\subsection{Multi Probe \lsh}
\label{sec:mblsh}
As we discussed in Section \ref{subsec:resuseHash}, generating hash functions can be computationally expensive.
Since the memory required by the algorithm also scales linearly with $L$, the number of hash tables, it is desirable to have a small 
number of tables to reduce the memory footprint. The memory footprint of vanilla LSH is what makes it impractical for real applications. 
Here, we first describe four new variants of the vanilla \lsh algorithm motivated by the intuition in
\mblshf~\cite{LvVLDB07}. \mblshf obtains significantly higher recall than the vanilla \lsh while using
the same number of hash tables. In order to achieve this, the main intuition utilized in \mblshf is that
in addition to looking at the hash bucket that a test query $q$ falls in, it is also possible to look at the neighboring buckets
in order to find its near neighbor candidates. \mblshf in ~\cite{LvVLDB07} suggests exploring the neighboring buckets in order of the
Hamming distance from the bucket in which $q$ falls. They then empirically show that these neighboring buckets contain 
the near neighbors with very high probability. 
Although \mblshf is achieves a higher recall for the same number of hash tables, it will also require more 
probes since it searches for multiple buckets within a table. 
The main advantage of searching in multiple buckets over generating more number of tables 
is that it takes more memory and time to generate more tables in pre-processing.
